Probabilistic Passphrase Cracking

### 2018-02-07 paper/paper.tex
Paper updated: added comparison table, and minor textual changes

Password cracking tools are widely available, which help us understand the
strength (or weakness) of passwords. For passphrases, there is no commonly used
approach using a single tool or dataset, probably because there are very few of
them.

@@

Sparell and Simovits\cite{sparell-simovits} created a passphrase cracking
program by using Markov chains on a word level. Markov chains are used in many
applications, such as next word prediction in keyboards on smartphones.
Similarly, one can use them to approximate the next most likely word in a
passphrase given a few starting words (which could be a dictionary walk). In
Sparell and Simovits' work, the LinkedIn hash dump of 2012 was used as one of
the test sets. This contains mostly passwords (not phrases), but is still
frequently used due to its large size of 64 million hashes. Of these, 21 000
phrases were cracked of between 10 and 20 characters in length.

@@

Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
dump (139 million MD5 entries) used a compilation of different public sources.
Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
or longer. Public sources include quotes from Wikiquote and text written in
bold and italics on Wikipedia.

Bonneau and Shutova\cite{payphrase-properties} examined ,,patterns of human
choice in a passphrase-level authentication system deployed by Amazon''. They
,,tested the availability of [over] 100,000 possible phrases at Amazon's
registration page, which prohibits using any phrase already registered by
another user.'' This system allows passphrases as short as two words and
offered no explanation to users regarding what might be secure. Its interface
merely asks for a "phrase" (which is generally a string of {\it logical} words)
and even suggests two-word phrases to use. The conclusion drawn by this paper
is not surprising, given these circumstances: ,,users aren't able to choose
phrases made of completely random words, but are influenced by the probability
of a phrase occurring in natural language. Examining the surprisingly weak
distribution of phrases in natural language, we can conclude that even 4-word
phrases probably provide less than 30 bits of security which is insufficient
against offline attack.'' We consider this low entropy estimation to be
specific for Amazon PayPhrase.

@@

In order to train a probabilistic algorithm, a dataset is needed to train on.
Labrande described cracking over 4 million entries from the
Korelogic\cite{korelogic} MD5 dataset, 200 000 of which are 16 or more
characters. This was done by downloading all articles of Wikipedia and
Wikiquote in various languages, plus a few other, smaller sources. Page titles,
list item entries, bold text, and italic text were used as phrases. Finally,
they used rules in John the Ripper to test variants of each phrase. They also
used rules which mangle the phrase into a password, for example by taking only
the first letter of each word in the phrase.

@@

Because we are mainly interested in English passphrases, we downloaded only the
English Wikipedia and Wikiquote articles. For a set of hashes, we used the
LinkedIn as well as the Korelogic dump. This also provides a way of comparing
results with Sparell and Simovits'\cite{sparell-simovits} Markov chain-based
method, which they tested on the LinkedIn set. We processed the data in a
similar manner: extracting text in italics and bold, page titles, and list
items on Wikiquote. We used only a few phrase mangling rules, which are
described in section \ref{sec:implementation}. Since we are, at this point,
only interested in passphrases, we did not include many of the more elaborate
rules: rules such as taking the first letter of each word or appending random
numbers, were not used. Such random variations would defeat the purpose of a
passphrase for the user, which is to be a set of common words with no
punctuation or other random tweaks to remember.

@@

Additionally, we have downloaded a large plain text password dump, specifically
the `breach compilation' of 1.4 billion passwords. This proved to mainly
consist of very insecure passwords and did not contain more than a few
passphrases.

@@

As far as we have been able to find, probabilistic context-free grammar has not
been applied to passphrase cracking. It proved to be a viable method in
Weir's\cite{weir} password cracking work. For passphrases, it could be adapted
to operate on whole words rather than individual characters. Based on texts and
previously cracked passphrases, rules could be derived which indicate for
example how likely it is for a noun to follow a verb.

@@

While analyzing the dataset, we saw that it contained many geographical names,
brands and person's names. Additionally, many of the phrases have the ring of a
catchphrase to them or form a common saying, for example "eye for an eye" or
"give me a break". While a probabilistic context-free grammar model might be
able to approximate those using large lists of names (and other words like
verbs, adjectives, etc.), n-grams capture parts of phrases and will inherently
become biased towards common phrases and constructions.

N-grams automatically include names and common constructions because they are
subsets of sentences, and implementations can be very fast because the
algorithm is fairly simple. The downside is that it has no memory: if it learns
`there are' and `are there' are common constructions, it will generate `are
there are', `there are there'. Markov chains have the same issue, but
context-free grammar does maintain more context (despite its name). We
chose to implement a proof of concept using n-grams because it is a novel
technique to generate passphrases and it seems fit for the purpose.

\section{Implementation}\label{sec:implementation}

@@

We implemented a program in Python which splits a text on spaces and counts
each possible subsentence of {\it n} words. For example, in the sentence `lay
of the land', trigrams are `lay of the' and `of the land'. The number of times
an n-gram occurs is counted. Another Python program uses these n-grams to
generate phrases, using the most frequently occurring n-grams first. The source
code for these programs can be found in our git repository\cite{git}. The
output of the second program can be directly used in tools such as Hashcat.

@@

A bigrams file was initially generated from the passphrases we cracked in
section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
because most users stripped their passphrases of spaces. We attempted to use
the program which we used in section \ref{sec:methods-comparison} to find
concatenated words within phrases, splitting e.g. `batterystaple' into
`battery' and `staple'. Again we found that the identification errors were too
great to be of any use, even when using an additional
dictionary\cite{wamerican} which contains conjugations. (This dictionary was
not used initially because it does not have word type classifications.)

In order to supplement the system with more possible sentence continuations, we
used English Wikipedia articles. This yielded around 29 million bigrams and 47
million trigrams. We combined them with the n-grams from the cracked
passphrases, prioritizing the latter. This way, n-grams from Wikipedia are only
used supplementarily and n-grams from previously cracked passphrases are
tried first.

Finally, we used mangling rules in Hashcat to create simple variants for each
phrase. Hashcat can use multiple rule sets which are then combined, for example
if in set 1 there are rules for both lowercasing and uppercasing a phrase and
in set 2 there are rules for removing spaces, the resulting phrases for `A b'
will be `a b', `A B', `ab' and `AB'. In our previous work, we learned what kind
of alterations users make to their phrases. Based on that information, we
created three sets of rules:

@@

The system we used for all tests is a laptop with an Intel i7-3630QM CPU and
8GiB RAM.

@@

In this section we will evaluate the performance of different passphrase
cracking methods, as well as our n-gram implementation. We consider the number
of entries they manage to crack as well as their efficiency.

Because we reproduced Labrande's hybrid dictionary attack (with slight
alterations) and cracked a set of hashes identical to those of Sparell and
Simovits'\cite{sparell-simovits}, we can compare the performance of all methods
directly.

@@

Of the LinkedIn set, we were able to crack 1.3 million entries, of which 13 361
of 16 characters or more and 26 345 cracked passwords include a space.

@@

The storage requirements of this method are relatively large: our phrases
file (with 26 million entries) is 690MiB. Since we had to generate this from
Wikipedia, we also had to download all Wikipedia and Wikiquote articles, which
are 26GiB as download and 135GiB uncompressed.

@@

Sparell and Simovits cracked 25 000 entries from the LinkedIn set, of which 384
of 16 characters or longer.

@@

Using our passphrase generator with trigrams, we were able to crack 835 562
entries of the Korelogic MD5 set, of which 32 911 are 16 characters or longer,
and 70 198 contain a space. Of the LinkedIn set, we were able to crack 482 014
entries, of which 3 782 are 16 characters or longer and 14 903 contain a space.

Using bigrams, we were able to crack 342 145 entries of the Korelogic set, of
which 9 148 are 16 characters or longer and 1 023 contain a space. Of the
LinkedIn set, we were able to crack 190 117 entries, of which 1 468 are 16
characters or longer and 1 449 contain a space.

@@

@@

We have not been able to explore probabilistic context-free grammar fully due
to the difficulties in classifying words. Weir was successful in this with
passwords because he focused on a character level. For passphrases, which are
composed of words instead of characters, this method is more complex. We
recommend looking into other, possibly commercial options for a word
classification database. We also do not recommend using previously cracked
passphrases as training set: many are stripped of spaces, making it very
difficult to programmatically split words for classification.

@@

\printbibliography

>>>>

Password cracking tools, which help us understand the strength (or weakness) of
passwords, are widely available. For passphrases, there is no commonly used
approach using a single tool or dataset, probably because there are very few of
them.

@@

Sparell and Simovits\cite{sparell-simovits} created a passphrase cracking
program by using Markov chains on a word level. Markov chains are used in many
applications, such as next word prediction in keyboards on smartphones.
Similarly, one can use them to approximate the next most likely word in a
passphrase given a few starting words (which could be a dictionary walk). In
Sparell and Simovits' work, the LinkedIn hash dump of 2012 was used as one of
the test sets. This contains mostly passwords (not phrases), but is still
frequently used due to its large size of 64 million hashes. Of these, 21 000
phrases of between 10 and 20 characters in length were cracked.

@@

Labrande\cite{crackmeimfamous} attempted to crack hashes from the Korelogic
dump (139 million MD5 entries) using a compilation of different public sources.
Around 4.3 million passwords were cracked, of which 200 000 are 16 characters
or longer. Public sources include quotes from Wikiquote and text written in
bold and italics on Wikipedia.

Bonneau and Shutova\cite{payphrase-properties} examined ,,patterns of human
choice in a passphrase-level authentication system deployed by Amazon''. They
,,tested the availability of [over] 100,000 possible phrases at Amazon's
registration page, which prohibits using any phrase already registered by
another user.'' This system allows passphrases as short as two words and
offered no explanation to users regarding what might be secure. Its interface
merely asks for a `phrase' (which is generally a string of {\it logical} words)
and even suggests two-word phrases to use. The conclusion drawn by this paper
is not surprising, given these circumstances: ,,users aren't able to choose
phrases made of completely random words, but are influenced by the probability
of a phrase occurring in natural language. Examining the surprisingly weak
distribution of phrases in natural language, we can conclude that even 4-word
phrases probably provide less than 30 bits of security which is insufficient
against offline attack.'' We consider this low entropy estimation to be
specific for Amazon PayPhrase.

@@

In order to train a probabilistic algorithm, a dataset is needed to train on.
Labrande described cracking over 4 million entries from the
Korelogic\cite{korelogic} MD5 dataset, 200 000 of which are 16 or more
characters. This was done by downloading all articles of Wikipedia and
Wikiquote in various languages, plus a few other, smaller sources. Page titles,
list item entries, bold text, and italic text were used as phrases. Finally,
they used rules in John the Ripper to test variants of each phrase. They also
used rules that mangle the phrase into a password, for example by taking only
the first letter of each word in the phrase.

@@

Since we are mainly interested in English passphrases, we downloaded only the
English Wikipedia and Wikiquote articles. For a set of hashes, we used the
LinkedIn as well as the Korelogic dump. This also provides a way of comparing
results with Sparell and Simovits'\cite{sparell-simovits} Markov chain-based
method, which they tested on the LinkedIn set. We processed the data in a
similar manner: extracting text in italics and bold, page titles, and list
items on Wikiquote. We used only a few phrase mangling rules, which are
described in section \ref{sec:implementation}. Since we are, at this point,
only interested in passphrases, we did not include many of the more elaborate
rules: rules such as taking the first letter of each word or appending random
numbers, were not used. Such random variations would defeat the purpose of a
passphrase for the user, which is to be a set of common words with no
punctuation or other random tweaks to remember.

@@

Additionally, a large plain text password dump was downloaded, specifically the
`breach compilation' of 1.4 billion passwords. This proved to mainly consist of
very insecure passwords and did not contain more than a few passphrases.

Note that the downloaded dumps did not contain usernames, and any cracked hash
values are kept confidential.

@@

So far, probabilistic context-free grammar has not been applied to passphrase
cracking. It proved to be a viable method in Weir's\cite{weir} password
cracking work. For passphrases, it could be adapted to operate on whole words
rather than individual characters. Based on texts and previously cracked
passphrases, rules could be derived which indicate for example how likely it is
for a noun to follow a verb.

@@

While analyzing the dataset, it was found that it contained many geographical
names, brands and people's names. Additionally, many of the phrases have the
ring of a catchphrase to them or form a common saying, for example "eye for an
eye" or "give me a break". While a probabilistic context-free grammar model
might be able to approximate those using large lists of names (and other words
like verbs, adjectives, etc.), n-grams capture parts of phrases and will
inherently become biased towards common phrases and constructions.

N-grams automatically include names and common constructions because they are
subsets of sentences, and implementations can be very fast due to the
simplicity of the algorithm. The downside is that it has no memory: if it
learns `there are' and `are there' are common constructions, it will generate
`are there are', `there are there'. Markov chains have the same issue, but
context-free grammar does maintain more context (despite its name). We chose to
implement a proof of concept using n-grams because it is a novel technique to
generate passphrases and it seems fit for the purpose.

\section{N-gram implementation}\label{sec:implementation}

@@

A program was implemented in Python which splits a text on spaces and counts
each possible subsentence of {\it n} words. For example, in the sentence `lay
of the land', trigrams are `lay of the' and `of the land'. The number of times
an n-gram occurs is counted. Another Python program uses these n-grams to
generate phrases, using the most frequently occurring n-grams first. The source
code for these programs can be found in our git repository\cite{git}. The
output of the second program can be directly used in tools such as Hashcat.

@@

A bigrams file was initially generated from the passphrases that were cracked
in section \ref{sec:dataset}. This yielded only 761 bigrams and 331 trigrams
because most users stripped their passphrases of spaces. We attempted to use
the program mentioned in section \ref{sec:methods-comparison} to find
concatenated words within phrases, splitting e.g. `batterystaple' into
`battery' and `staple'. Again we found that the identification errors were too
great to be of any use, even when using an additional
dictionary\cite{wamerican} which contains conjugations. (This dictionary was
not used initially because it does not have word type classifications.)

In order to supplement the system with more possible sentence continuations, we
used English Wikipedia articles. This yielded around 29 million bigrams and 47
million trigrams. Those were combined with the n-grams from the cracked
passphrases, prioritizing the latter. This way, n-grams from Wikipedia are only
used supplementarily and n-grams from previously cracked passphrases are tried
first.

Finally, mangling rules were used in Hashcat to create simple variants for each
phrase. Hashcat can use multiple rule sets which are then combined. If, for
example, if in set 1 there are rules for both lowercasing and uppercasing a
phrase and in set 2 there are rules for removing spaces, the resulting phrases
for `A b' will be `a b', `A B', `ab' and `AB'. In our previous work, we learned
what kind of alterations users make to their phrases. Based on that information,
we created three sets of rules:

@@

The system used for all tests is a laptop with an Intel i7-3630QM CPU and 8GiB
RAM.

@@

In this section the performance of different passphrase cracking methods is
evaluated, considering both the number of entries they manage to crack and
their efficiency.

Because we reproduced Labrande's hybrid dictionary attack (with slight
alterations) and cracked a set of hashes identical to those of Sparell and
Simovits'\cite{sparell-simovits}, the performance of all methods can be
compared directly.

@@

Of the LinkedIn set, we were able to crack 1.3 million entries, of which 13 361
are 16 characters or more and 26 345 cracked passwords include a space.

@@

The storage requirements of this method are relatively large: our phrases file
(with 26 million entries) is 690MiB. Since this was generated from Wikipedia,
all Wikipedia and Wikiquote articles also had to be downloaded, which total
26GiB as download and 135GiB uncompressed.

@@

Sparell and Simovits cracked 25 000 entries from the LinkedIn set, of which 384
are 16 characters or longer.

@@

Using the passphrase generator with trigrams, it was able to crack 835 562
entries of the Korelogic MD5 set, of which 32 911 are 16 characters or longer,
and 70 198 contain a space. Of the LinkedIn set, it was able to crack 482 014
entries, of which 3 782 are 16 characters or longer and 14 903 contain a space.

Using bigrams, it was able to crack 342 145 entries of the Korelogic set, of
which 9 148 are 16 characters or longer and 1 023 contain a space. Of the
LinkedIn set, it was able to crack 190 117 entries, of which 1 468 are 16
characters or longer and 1 449 contain a space.

@@

\subsection{Overview}

Table \ref{tab:comparison} gives an overview of the performance of the
different methods. The column headers `Dict-1' and `Dict-2' indicate the hybrid
dictionary attack, Labrande's and ours, respectively. The rows labeled `16+'
indicate the number of results which are 16 characters or longer.

Note that for our hybrid dictionary attack, the number of generated phrases per
second is estimated at 10 million. Contrarily to all other methods, this method
is IO-bound instead of CPU-bound. The amount of RAM required is considered
negligible, as it only requires IO buffers.

\begin{table}[H]
	\begin{tabular}{| l | r | r | r | r | }
	\hline                & {\bf Dict-1} & {\bf Dict-2} & {\bf Markov} & {\bf 3-grams} \\
	\hline LinkedIn       & n/a          &  1 300 000   &    25 000    &   482 014     \\
	\hline LinkedIn 16+   & n/a          &     13 361   &       384    &     3 782     \\
	\hline Korelogic      & 4 200 000    &  2 300 000   & n/a          &   835 562     \\
	\hline Korelogic 16+  &   200 000    &    146 742   & n/a          &    32 911     \\
	\hline Phrases/second & n/a          & 10 000 000   &    22 500    & 3 300 000     \\
	\hline RAM            & n/a          & negligible   & n/a          & 5.3GiB        \\
	\hline Storage        & n/a          &     690MiB   & n/a          & 464MiB        \\
	\hline
	\end{tabular}
	\caption{Performance overview}
	\label{tab:comparison}
\end{table}

@@

We have not been able to explore probabilistic context-free grammar fully due
to the difficulties in classifying words. Weir was successful in this with
passwords because he focused on a character level. For passphrases, which are
composed of words instead of characters, this method is more complex. It is
recommended to look into other, possibly commercial options for a word
classification database. It is also recommended to avoid using previously
cracked passphrases as training set: many are stripped of spaces, making it
very difficult to programmatically split words for classification.

@@

\newpage
\printbibliography
